Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-10-21 manuscript_latex/manuscript.tex
Fixed spelling errors.

In simulations and real-world datasets, robust location- and scale-invariant 
transformations outperform conventional variants, resulting in better 
transformations to central normality. In a machine learning experiment
with 232 datasets with numerical features, integrating robust 
ocation- and scale-invariant power transformations into an automated data 
processing and machine learning pipeline did not result in a meaningful
improvement or detriment in model performance compared to conventional variants.

@@

\begin{highlights}
\item Location and scale of feature distributions may cause wrong power transformations.
\item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
\item Robust power transformations succesfully reduced negative effects of outliers.
\item New invariant robust power transformations can replace conventional variants.
\end{highlights}

@@

\caption{Weighting functions investigated in this study to make power transformations more robust against outliers. In this example, the step function was parameterised with $\delta_1 = 0.60$. The triangle and tapered cosine functions were both parameterised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
\end{figure}

@@

Location- and scale-invariant power transformations are intended to
yield improved transformations to normality in the presence of large
shifts in location, distributions that due to location and scale are not
centered near zero, or both. Earlier, we assessed these transformations
using simulated data. In the following, they are evaluated using
examples from real datasets. We focus on the Yeo-Johnson transformation
because of its ability to handle features with negative values. Results
for Box-Cox transformations are shown in 
\ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.

@@

\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets. The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). Random forests outperform GLM models for all task difficulties. The top-right panel shows the marginal effect of feature normalisation methods, i.e. no normalisation and z-standardisation. z-standardisation is generally beneficial, but the estimated effect is negligible. The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method. There is no consistent behaviour, and estimated effects are negligible. Note that the ranges of the $x$-axes of the three main panels differ. ECNT: empirical central normality test.}\label{fig:marginal-effect-plot}
\end{figure}

>>>>

In simulations and real-world datasets, robust location- and scale-invariant 
transformations outperform conventional variants, resulting in better 
transformations to central normality. In a machine learning experiment
with 232 datasets with numerical features, integrating robust 
location- and scale-invariant power transformations into an automated data 
processing and machine learning pipeline did not result in a meaningful
improvement or detriment in model performance compared to conventional variants.

@@

\begin{highlights}
\item Location and scale of feature distributions may cause wrong power transformations.
\item We derived location- and scale-invariant Box-Cox and Yeo-Johnson transformations.
\item Robust power transformations successfully reduced negative effects of outliers.
\item New invariant robust power transformations can replace conventional variants.
\end{highlights}

@@

\caption{Weighting functions investigated in this study to make power transformations more robust against outliers. In this example, the step function was parametrised with $\delta_1 = 0.60$. The triangle and tapered cosine functions were both parametrised with $\delta_1 = 0.30$ and $\delta_2 = 0.90$.}\label{fig:weighting-functions}
\end{figure}

@@

Location- and scale-invariant power transformations are intended to
yield improved transformations to normality in the presence of large
shifts in location, distributions that due to location and scale are not
centred near zero, or both. Earlier, we assessed these transformations
using simulated data. In the following, they are evaluated using
examples from real datasets. We focus on the Yeo-Johnson transformation
because of its ability to handle features with negative values. Results
for Box-Cox transformations are shown in 
\ref{appendix-d-experimental-results-using-location--and-scale-invariant-box-cox-transformation}.

@@

\caption{Estimated marginal effect of learners, normalisation and transformation methods on ranked model performance scores in 18560 machine learning experiments on 232 datasets. The top-left panel shows the marginal effect of learners, i.e. random forests and generalised linear models (GLM). Random forests outperform GLMs for all task difficulties. The top-right panel shows the marginal effect of feature normalisation methods, i.e. no normalisation and z-standardisation. z-standardisation is generally beneficial, but the estimated effect is negligible. The bottom panel shows the marginal effects of different transformation methods, split by learner and normalisation method. There is no consistent behaviour, and estimated effects are negligible. Note that the ranges of the $x$-axes of the three main panels differ. ECNT: empirical central normality test.}\label{fig:marginal-effect-plot}
\end{figure}
